---
title: "MATH 216 Homework 3"
author: "Bianca Gonzalez"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(readr)
library(lubridate)
library(Quandl)
library(forcats)
library(knitr)
```

## Admistrative:
Please indicate

* Who you collaborated with:
Brenda, Shannia, Albert Kim. 

* Roughly how much time you spent on this HW so far:
5 hours
* The URL of the RPubs published URL [here](http://rpubs.com/biancaglez94/223626).
* What gave you the most trouble:
the training dataset and lubridate. 
* Any comments you have:
Thanks!

## Question 1:


We will use a logistic regression model to predict sex. Our metric to rate how well our
model performs will be:

$$
\frac{1}{n}\sum_{i=1}^{n}I(y_i = \widehat{y}_i)
$$

where $I(A)$ is the *indicator function* that is equal to 1 if condition $A$
holds, 0 otherwise. So

* Say user $i$ **is** female, then $y_i=1$
* Say we **predict** user $i$ is female, then $\widehat{y}_i=1$
* In this case $I(y_i =\widehat{y}_i)=1$. 

So what the above formula is reporting is the proportion of users' sex we
correctly predicted.

```{r, echo=FALSE, message=FALSE, cache=TRUE}
# Edit this code block at your own peril! cache is set to TRUE!
# To keep this exercise simple, let's remove the only 3 users (0.005% of users)
# who did not list a height, define the outcome variable, and add an ID variable
# to distinguish the users

profiles <- read_csv(file="/Users/BiancaGonzalez/Desktop/RudeBoyBert/HW-02/profiles.csv") %>% 
  filter(!is.na(height)) %>% 
  mutate(is_female=ifelse(sex=='f', 1, 0)) %>% 
  tibble::rownames_to_column(var="id")
set.seed(76)
```



#### a)

```{r, echo=FALSE, message=FALSE, cache=TRUE}
#This is the other categorical variable I considered. 

profiles <- profiles %>% 
  mutate(diet = ifelse(is.na(diet), "did not report", diet))

profiles <- profiles %>% 
  mutate(job = ifelse(is.na(job), "did not report", job))

profiles <- profiles %>%
  mutate(diet_code = fct_recode(diet,
                            # New name = old name
                            "omnivore" = "mostly anything",
                            "omnivore" = "anything",
                            "omnivore" = "strictly anything",
                            "omnivore" = "mostly other",
                            "omnivore" = "strictly other",
                            "omnivore" = "other",
                            "veggie" = "mostly vegetarian",
                            "veggie" = "strictly vegetarian",
                            "veggie" = "mostly vegan",
                            "veggie" = "vegetarian",
                            "veggie" = "strictly vegan",
                            "veggie" = "vegan",
                            "other" = "mostly kosher",
                            "other" = "mostly halal",
                            "other" = "strictly kosher",
                            "other" = "strictly halal",
                            "other" = "halal",
                            "other" = "kosher",
                            "did not report" = "did not report"
  )) 


```


Define:

* A *training* set `training` of 2997 users (5% of users). We will train the 
logistic regression model to predict gender using this data. Since we want to 
train the model to tell who is female and who is not, we use the outcome
variable `is_female`.

* A *test* set `test` of the remaining 56,946 users (95% of users). We will test
how good our trained model is using this data. 

So at first, we will pretend we
don't know the outcome variable `is_female`. We use the above model to make a
prediction of sex for all 56,946 test users, then we use the `is_female` outcome
to rate how well we performed.
* Be sure to incorporate all the insight your garnered in your EDA in HW-2.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}

profiles <- profiles %>%
  mutate(is_female = ifelse(sex=="f", 1, 0)) %>%
  mutate(
    last_online = stringr::str_sub(last_online, 1, 10),
    last_online = lubridate::ymd(last_online)
  )

#a training set of 5% of users:
train5 <- profiles %>% 
  sample_frac(0.05)

#testing set with 95% of users
test95 <- anti_join(profiles, train5, by="id")

```

#### b)

Train the logistic regression model to predict sex. i.e. fit a logistic
regression model to the `training` data. Assign this model to an R object called
`predict_sex_model`, then rate how well the model performs on the `training` data.

```{r, echo=FALSE, message=FALSE, cache=TRUE}

predict_sex_model <- glm(is_female ~ income + job + diet_code, data=train5, family="binomial")

#broom::tidy(predict_sex_model)

```


#### c)

Take `predict_sex_model` and apply it to the `test` data and make a prediction 
for each users' sex, then rate how well the model performs on the `test` data.

**Hint**: What do you think `predict(predict_sex_model, newdata=test,
type="response")` does? The help file is located in `?predict.glm`


The first output is the output for the training data that the model was built on. This is 63% accurate. My test data I ran the model on was 61% correct. 
```{r, echo=FALSE, message=FALSE, cache=TRUE}

#this will give us our phats which we can then create a threshold for. 
#type = response gives us probabilities. 

#phat not in predictsexmodel generation

#we are testing the model we built with the training data by finding 
#pHats (probability that is female)! 
test95 <- test95 %>% 
  mutate(p_hat = predict(predict_sex_model, newdata=test95, type="response"))

test95 <- test95 %>% 
  select(income, job, diet_code, p_hat, is_female)

#already used train data to build the model so this is just getting our error from our training data. 
train5 <- train5 %>% 
  mutate(p_hat = predict(predict_sex_model, newdata=train5, type="response"))


train5 <- train5 %>% 
  mutate(threshold_prediction = ifelse(p_hat >.6, 1, 0)) %>% 
  mutate(correct_prediction = ifelse(is_female==threshold_prediction, 1, 0)) 

mean(train5$correct_prediction)

#threshold to evaluate if correct. phat. use mutate to create new phat column
#where you put the generated responses from predictions. 


#ifelse to see if the phat is in threshold, if it is within threshold, then mark as female. then can compare both rows and see if they are correct. 

test95 <- test95 %>% 
  mutate(threshold_prediction = ifelse(p_hat >.6, 1, 0)) %>% 
  mutate(correct_prediction = ifelse(is_female==threshold_prediction, 1, 0)) 

mean(test95$correct_prediction)


#now have to compare rate of accuracy by grouby true/false and then divide by total to see
#percent accurate. 
#count(is_female==is_female_predic) %>%  mutate(prop=m/sum(n))
                           
```

#### d)

Did the model perform better on the `training` data or the `test` data? Why
do you think that is?

My training model was 2% more accurate than my test data analysis. I think this is the case because the model was built on the training data set. 


## Question 2:

We want to compare the volatility of 

* [Bitcoin](https://www.quandl.com/data/BAVERAGE/USD) prices
* [Gold](https://www.quandl.com/data/BUNDESBANK/BBK01_WT5511) prices

Let our measure of volatility be the relative change from day-to-day in price. 
Let the reference currency be US dollars. Analyze these results and provide
insight to a foreign currency exchanger.

```{r, results = 'asis', echo=FALSE}

#find difference between days. Absolute change
#lead is command for lag.... 
#finding absolute change over days. 
#bindrows. puts underneath
#same column structure. select date, value for both, and create a column called type.(mutate)
#groupby_type..... 
# 1. Get gold & bitcoin data
# 2. Make column structure the same
# 3. Add variable type

#have to explore the data prior so you can find the variables you can select or rename
gold <- Quandl("BUNDESBANK/BBK01_WT5511") %>% 
  select(Date, Value) %>%             #Here you make the data the same so you can bind the rows                                             together. 
  mutate(type="Gold")

bitcoin <- Quandl("BAVERAGE/USD") %>% 
  rename(Value = `24h Average`) %>%       #rename the variable to be the same as the other 
  select(Date, Value) %>% 
  mutate(type="Bitcoin")

# Combine them into single data frame using bind_rows()
combo_data <- bind_rows(gold, bitcoin) %>% 
  # Group by here!
  group_by(type) %>% 
  # Then do the following ONLY ONCE:
  filter(year(Date) >= 2011) %>%        #find year where gold begins prior to combining data
  arrange(Date) %>%                     #arrange them by date and check out order b4 using function.
  mutate(
    Value_yest = lag(Value),            #create a lagged variable so you can subtract diff./divde
    rel_diff = 100 * (Value-Value_yest)/Value_yest    #calculate percentage using this new var
    )

# Plot
ggplot(combo_data, aes(x=Date, y=rel_diff, col=type)) +
  geom_line() +
  labs(title = "Bitcoin and Gold Volatitily", y="% Change")

#relative change from day to day in price of bitcoin and gold. 
#so need to be able to create dataset where dates match, then subtract gold from bitcoin 
#and find the absolute value to show the change in these two.

#labs(Title = "Absolute change in price of Gold and bitcoin from 2011 to 2016")



```


## Question 3:

```{r, echo=FALSE, message=FALSE, cache=TRUE}
# Edit this code block at your own peril! cache is set to TRUE!
jukebox <- read_csv(file="reed_jukebox.csv")

# Clean certain artists' names:
sigur_ros <- "Sigur Ro\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8c\x93\xa0\xbcs"
bjork <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk"
blue_oyster_cult <- "Blue O\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcyster Cult"
husker_do <- "Hu\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcsker Du\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbc"
bjork_brodsky <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk & Brodsky Quartet"
slagsmalsklubben <- "Slagsma\xfc\xbe\x99\x86\x94\xbc_lsklubben "
bjork_sugarcubes <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk (Sugarcubes)"

jukebox <- jukebox %>%
  mutate(
    artist = ifelse(artist == sigur_ros, "Sigor Ros", artist),
    artist = ifelse(artist == bjork, "Bjork", artist),
    artist = ifelse(artist == blue_oyster_cult, "Blue Oyster Cult", artist),
    artist = ifelse(artist == husker_do, "Husker Do", artist),
    artist = ifelse(artist == bjork_brodsky, "Bjork & Brodsky Quartet", artist),
    artist = ifelse(artist == slagsmalsklubben, "Slagsmalsklubben", artist),
    artist = ifelse(artist == bjork_sugarcubes, "Bjork (Sugarcubes)", artist)
  )
```


Using the Reed College jukebox data, what are the top 10 artists played during
the "graveyard shift" during the academic year? Define

* the "graveyard shift" as midnight to 8am
* the academic year as September through May (inclusive)

```{r, results = 'asis', echo=FALSE}
#jukebox
#jukebox$date_time %>% 
 # class()
#of type character so let's convert this baby to a date format. 
#had to add timezone. UTC!
#first let's edit the times before continuing . 

#ALSO EDITED TIME ZONE!
jukebox <- jukebox %>% 
  mutate(
    date_time = parse_date_time(date_time, "a b d HMS Y"),
    date_time = with_tz(date_time, tz = "America/Los_Angeles")
  )
#?as.factor()   this converts to a categorical variable 

juke <- jukebox %>% 
  filter(month(date_time) %in% c(9,5,10,11,12,1,2,3,4) == TRUE) %>% 
  filter(hour(date_time) %in% c(00, 01, 02,03,04,05,06,07,08) == TRUE)
    
 
juke_ten <- juke %>% 
  group_by(artist) %>% 
  tally() %>% 
  arrange(desc(n)) %>% 
  slice(1:10) %>% 
kable()

juke_ten

#then find top ten artists (max) (asc) kable, splice! show ten! 

  


```





